Paper Title*

### 2018-12-24 paper.tex
intro intro

\section{Introduction}
\par In the last few years, number of Internet of Things applications increase exponentially. \cite{7721743} Recent developments on LPWAN technologies has great impact on IoT application number growth. LPWAN technologies offer solutions to some of the oldest wireless communication challenges. Traditional wireless communication methods such as cellular networks (e.g., 2G, 3G, LTE) and short-range communication methods (e.g., Bluetooth, WiFi, Zigbee) cannot provide low power and long range at the same time. Cellular networks can provide long range and high data rate but they are complex and consume too much power. Besides, most of the IoT applications don't require high data rate. Short-range communication methods can provide relatively low power consumption but their range is limited to a few hundred meters at best. \cite{7815384} LPWAN technologies fill the technology gap between short range and cellular communication by providing low power and long range communication. LPWAN technologies basically sacrifice data rate to provide low power consumption.

\par There are several emerging LPWAN technologies. LoRa, SigFox, NB-IoT and LTE-M are commonly used, well known LPWAN technologies. LoRa and SigFox use license free ISM frequency bands while NB-IoT and LTE-M use licensed frequency bands which brings extra cost. Both LoRa and SigFox known for ultra low power consumption and resilient to interference while NB-IoT and LTE-M are promoted for higher data rate.

LoRa has open standard MAC protocol called LoRaWAN.\\
LoRaWAN and SigFox use unslotted ALOHA based MAC protocol.\\
Private LoRaWAN networks can be deployed however SigFox and NB-IoT are operator based.\\
Sigfox limits the number of messages an end device can send (140 UL, 4 DL per day).\\
Sigfox payload is very limited 12 bytes UL and 8 bytes DL. LoRa 243 bytes, NB-IoT 1600 bytes.\\
Sigfox maximum data rate is also very low 100 bps, LoRa 50 kbps, NB-IoT 200 kbps.\\
LoRa can change data rate by spreading symbols within a fixed channel bandwidth. This enables tradeoff between sensitivity and data rate. LoRa has 7 different spreading factor option. Different spreading factor communications are orthogonal to each other.

@@

\subsection{LoRaWAN}
\par LoRaWAN is a media access control (MAC) layer protocol which designed for large scale LoRa networks. LoRaWAN is an open source standard developed and maintained by LoRa Alliance. LoRa Alliance is an open, non-profit organization dedicated to promoting the interoperability and standardization of LoRaWAN. LoRa can be used without LoRaWAN as a wireless link technology, however LoRaWAN is developed considering well known LPWAN challenges and their best practice solutions. LoRaWAN also provides interoperability between different networks. LoRaWAN uses unslotted ALOHA to access medium. LoRaWAN provides lightweight but powerful standard for IoT applications.

@@

\par In \cite{s17061193}, the authors investigated single gateway LoRaWAN network scalability in terms of the number of end nodes using a simulation model based on real measurements. They measure the impact of two concurrent LoRa transmissions on each other by using physical LoRaWAN end devices and a gateway then they created a simulation model from measurements. Their results show that LoRaWAN has better scalability than pure Aloha since a LoRa packet may still go through under collision if the the last six symbols of preamble and header of the packet does not collide.

>>>>

\section{Introduction}
\par In the last few years, number of Internet of Things applications increase exponentially. \cite{7721743} Recent developments on LPWAN technologies has great impact on IoT application number growth. LPWAN technologies offer solutions to some of the oldest wireless communication challenges. Traditional wireless communication methods such as cellular networks (e.g., 2G, 3G, LTE) and short-range communication methods (e.g., Bluetooth, WiFi, Zigbee) cannot provide low power and long range at the same time. (CITE) Cellular networks can provide long range and high data rate but they are complex and consume too much power. Besides, most of the IoT applications don't require high data rate. Short-range communication methods can provide relatively low power consumption but their range is limited to a few hundred meters at best. \cite{7815384} LPWAN technologies fill the technology gap between short range and cellular communication by providing low power and long range communication. LPWAN technologies basically sacrifice data rate to provide low power consumption.

\par There are several emerging LPWAN technologies. LoRa, SigFox, NB-IoT and LTE-M are commonly used, well known LPWAN technologies. LoRa and SigFox use license free ISM frequency bands while NB-IoT and LTE-M use licensed frequency bands which brings extra cost. (CITE) Both LoRa and SigFox known for ultra low power consumption and resilient to interference while NB-IoT and LTE-M are promoted for higher data rate. LoRa has open standard MAC protocol called LoRaWAN. (CITE) LoRaWAN and SigFox MAC protocols are based on pure ALOHA medium access. (CITE) LoRaWAN networks can be deployed as private network. However, SigFox and NB-IoT supports only operator contracted deployments. (CITE) Number of messages that Sigfox end device can send in a day is limited to 140 packets for uplink and just 4 packets for downlink. Sigfox packet payload is limited to 12 bytes for uplink and 8 bytes for downlink. (CITE) However LoRa supports up to 243 bytes payload and NB-IoT supports up to 1600 bytes payload. (CITE)
Sigfox maximum data rate is also very low 100 bps, LoRa 50 kbps, NB-IoT 200 kbps.\\
LoRa can change data rate by spreading symbols within a fixed channel bandwidth. This enables tradeoff between sensitivity and data rate. LoRa has 7 different spreading factor option. Different spreading factor communications are orthogonal to each other.

@@

\subsection{LoRaWAN}
\par LoRaWAN is a media access control (MAC) layer protocol which designed for large scale LoRa networks. LoRaWAN is an open source standard developed and maintained by LoRa Alliance. LoRa Alliance is an open, non-profit organization dedicated to promoting the interoperability and standardization of LoRaWAN. LoRa can be used without LoRaWAN as a wireless link technology, however LoRaWAN is developed considering well known LPWAN challenges and their best practice solutions. LoRaWAN also provides interoperability between different networks. LoRaWAN uses pure ALOHA to access medium. LoRaWAN provides lightweight but powerful standard for IoT applications.

@@

\par In \cite{s17061193}, the authors investigated single gateway LoRaWAN network scalability in terms of the number of end nodes using a simulation model based on real measurements. They measure the impact of two concurrent LoRa transmissions on each other by using physical LoRaWAN end devices and a gateway then they created a simulation model from measurements. Their results show that LoRaWAN has better scalability than pure ALOHA since a LoRa packet may still go through under collision if the the last six symbols of preamble and header of the packet does not collide.
